\section{INTRODUCTION}

\begin{flushright}
  \small{
  \textit{``The history of building construction can be construed as a narrative of the inertia and momentum of two divergent construction logistics. One mode[, discussed above,] has very minimal historical inertia coupled with great current industrial momentum (the muli-layered assemblies of modernity.) The other has great historical, physical, and thermodynamic inertia that is coupled with minimal industrial momentum in the contemporary building industry/building science industry (more monolithic assemblies and masses). The former follows the short history of the twentieth century ``rationalization" of construction, air-conditioning, factory production, lightweight envelopes, and, more recently, mass customization. The latter is a several-thousand-year history of accumulative knowledge and performance all but forgotten in the interesting yet hubristically selective amnesia of twentieth century architecture."}}\\ --- Kiel Moe. \\ \textit{Convergence}. 2013.
\end{flushright}

``Rammed-earth" refers to an earthen building material formed by a particular mechanical process. Rammed-earth architecture has ancient archaeological, anonymous, and autochthonous roots in China, Africa, Europe, India, the Middle East, and other regions globally \cite{CHRONO}. Observably, rammed-earth forms have been (re)appearing in the U.S. over the past half-century with a frequency and technical gain atypical of earthen construction, especially in the West. Contemporary rammed-earth forms are predominantly connected to academia, professional architectural design, and building science/industry.

Rammed-earth structures have appeared relatively recently on the campuses of M.I.T. (2005), Stanford (2015), and Princeton (2016)\footnote{\url{https://archive.is/5gPbZ} (M.I.T.); \url{https://archive.is/VhpW2} (Stanford); \url{https://archive.is/9SF6K} (Princeton)}. Rammed Earth Works (the designers of Stanford's Windhover Contemplative Center) is one of multiple professional architectural firms to be designing modernized rammed-earth buildings and installations valued in the multimillion-dollar range\footnote{\url{https://archive.is/K853p}}. David Easton, the founder of Rammed Earth Works (1976), also invented PISE (Pneumatically Impacted Rammed Earth) and Watershed Blocks (under a \$750k grant from the N.S.F.); a sprayed application of wet soil-cement and a mechanical system for mass-producing modular soil-cement blocks, respectively. SIREWALL (Structural Insulated Rammed-Earth, B.C., Canada) is a rammed-earth-based building product incorporating an intermediary layer of patented insulation\footnote{\url{https://archive.is/Sf9fu} (PISE); \url{https://archive.is/x3iwg} (Watershed Materials); \url{https://archive.is/s0cHI} (SIREWALL)}. Numerous technical papers concerning rammed-earth's structural and thermal properties have been published globally. A select few include: \textit{Analysis of the hygrothermal functional properties of stabilised rammed-earth materials} by Hall and Allinson. (2009), \textit{Modeling rammed earth wall using discrete element method} by Bui et al. (2015), and \textit{Measured and simulated thermal behaviour in rammed earth houses in a hot-arid climate. Part B: Comfort} by Beckett et al. (2017).

\vspace{5mm}

\underline{Hypothesis:} Along the decades of rammed-earth's previous resurgence in the U.S., at the eco-movements of the 60s and 70s, rammed-earth took on a greater socio-technical momentum while standards, practices, and policies of the building culture remained inert. The result was a form-based function of modern construction---a virtual image of sustainability---reducing, reusing, and recycling a historically function-based form of complexity, persistence, and adaptability. The rammed-earth material and method morphed more significantly in forty years than it had in four thousand.

If Doctor E. is right, and we truly can not solve problems by using the same kind of thinking we used when we created them, then it stands to reason that continued control of rammed-earth technology is not as much a determinant of sustainability as the model by which rammed-earth building is conceptualized, codified, communicated, designed, logisticized, and constructed.

Archaeological building principles and heuristics characterize sustainability as a pulsing journey, not an arrival at absolute steady-state.

\begin{flushright}
\small{
\textit{
``Sustainability is never a static goal. It can only be a process. Previous ideas about ``sustainability" are not and will never be tenable. A small, beautiful, modest, hand-crafted society, living in harmony with its eco-region, relentlessly parsimonious in its use of energy and resources, can't learn enough about itself to survive. In its bucolic quietude, it may appear timeless, but the clock is ticking for it as it does for all societies. It can avoid many conventional threats by abjuring large-scale, clumsy technologies, but modesty does not make one invisible. That society isn’t keeping track---in its loathing for industrialism, it forfeits far too much command-and-control over its physical circumstances. Its bliss is ignorance.\\
A truly sustainable society has to be sustainable enough to prevail against the unforeseen. The unforeseen, by definition, can’t be outplanned. This implies that serendipity is necessary. We can’t know what we need to know; so there need to be large stores of unplanned knowledge. . . The ability to make many small mistakes in a hurry is a vital accomplishment for any society that intends to be sustainable."}} \\ --- Bruce Sterling. \\ \textit{Shaping Things}. 2005.
\end{flushright}

\vspace{5mm}

\underline{Objective:} \textbf{Design a system} of design (a coalescing model of models) drawing from contemporary building theory and technology, reaching towards the principles and heuristics of the traditional rammed-earth material and method. The model is directed towards a computational system (digital accumulator and distributor of information) capable of organizing two flows:

1. The flow of soil types from deposit to building site (determinants of rammed-earth's embodied energy and building performance).

2. The flow of knowledge between builders (determinant of rammed-earth standardization and design).


\begin{flushright}
\small{
\textit{``[T]he culture that once was slow-moving, and allowed ample time for adaptation, now changes so rapidly that adaptation cannot keep up with it. No sooner is adjustment of one kind begun than the culture takes a further turn and forces the adjustment in a new direction. No adjustment is ever finished. And the essential condition on the process --- that it should in fact have time to reach its equilibrium --- is violated. This has all actually happened. In our own civilization, the process of adaptation and selection which we have seen at work in the unselfconscious cultures has plainly disappeared."}}\\ --- Christopher Alexander. \\ \textit{Notes on the Synthesis of Form}. 1964.
\end{flushright}


% A precedent similar in its regard for the past is the \textit{Manifesto for responsible architecture}, a one-page document assembled by a collective of architectural collectives and presented at the U.N. Climate Change Conference\in Paris (2015)\footnote{\url{https://web.archive.org/web/20180429205023/https://www.ace-cae.eu/fileadmin/New_Upload/8._Images/News/2015/Manifesto_EN_2.pdf}}. Allegedly, the Manifesto derived from the fossil of Caral, Peru (3,000-1800 B.C.), during a retrospective on the ``high engineering" involved in its structure, from ductwork to urban planning\footnote{\url{http://archive.is/e3wBp}}.



\clearpage

\subsection{The Model is the Message}

Marshall McLuhan noted (\textit{Understanding Media}, 1964) that, with respect to media/technology, the medium is the message. The rammed-earth building medium \textit{qua} pre-modern rammed-earth conveys a perennial socio-technical desire for a functional, durable, and economical form of building. Without any positive rammed-earth heritage, the U.S. has nonetheless considered or adopted rammed-earth construction during energy-sensitive phrases of its history. For instance, in the late-eighteenth century, French architect/builder Fran\c cois Cointereaux presented Thomas Jefferson and America's burgeoning rural economy with a model of rammed-earth architecture. Encoded in a copy of \textit{Ecole d'architecture rurale} (Paris, 1790-91); Cointereaux believed that if America adopted ``the economical building art of the ancients, perfected and made more universal," She would incur a great physiocratic power. Jefferson reacted indifferently, in a letter to William Short, ``how far it may offer benefit here superior to the methods of the country, founded in the actual circumstances of the country as to the combined costs of labour \& materials, and the circumstances of durability comfort \& appearance, must be the result of calculation."\footnote{\url{http://archive.is/yWexi} (Cointereaux to Jefferson); \url{http://archive.is/ozqQv} (Jefferson to Short)} Ostensibly, it would not be too long before industrialism captivated American hearts and minds and rammed-earth fell by the wayside.

Rammed-earth later appeared in \textit{Popular Mechanics (Vol. 41, No. 2, 1924)} and \textit{The Farmers Bulletin  (No. 1500, 1926)}, endorsed as a frugal, do-it-yourself building method. During the Great Depression, rammed-earth briefly held the attention of the New Deal-era Resettlement Administration as an economical building alternative fit for an over-abundance of available labor. During the 1960s and 1970s, rammed-earth attracted marginal interest from the environmental movements following the global recognition of troubling anthropogenic effects on the biosphere \cite{GARDENDALE}. Hassan Fathy wrote \textit{Architecture for the Poor} in 1973; painting earthen architecture as a bottom-up, practical way of rebuilding the Egyptian countryside. [Fathy spoke to a building culture that does possess a positive heritage in earthen architecture, opposed to the U.S. which possesses no such positive heritage.] Through these last resurgences of rammed-earth, the murmur of the D.I.Y. spirit dwindled and dissipated. \textit{Although the material would continue, the ``open" method was lost to a certain propriety of building practice.}

In a gross linearization of history, it would appear that a growing field around ``ecodevelopment" in the 1960s/1970s reintroduced rammed-earth building into a technologically dependent world caught between (in binary terms) developmentalist and zero-growth strategies\footnote{\url{http://archive.is/sOf7w}} to face the pertinent biospheric uncertainties. Hypothetically, at this shearing of [simply] technological optimism and technological pessimism, building societies retrofit rammed-earth into a hybrid model where cake could be had and also eaten. On one hand, the tried method would remain a virtual image of sustainability. On the other, seemingly innocuous technical changes to rammed-earth's composition and construction would modernize the material-method, ensuring scalability, standardization, and security.

Rammed-earth v2.0 manifested as ``soil-cement", also known as ``cement stabilized rammed-earth." Over time, it was layered with modernizing assemblages of mechanization, pre-fabrication, insulation, transportation, seal-ification, svelte-ification and modularization. An early code of practice was \textit{Soil-Cement: Its Use in Building}, distributed by the U.N. Department of Economic and Social Affairs in 1964.

\begin{flushright}
\small{
\textit{``The use of simple compacted soil (natural earth) as a building material dates from time immemorial, and it can be said that ever since, and down to the present day, the method of building houses with earth has been used, because of its constructive qualities. Yet, despite its good insulating and resistant properties, there are limitations to the use of earth owing to its lack of strength and its vulnerability to moisture and the erosive effects of external agents. Provided that natural soil possesses a combination of certain characteristics, however, it can be subjected to the process known as `stabilization'. The effect of adding a stabilizing agent like Portland cement, for instance, is not only to enhance its best qualities but to impart to it other properties which soil alone does not possess."}}\\ --- Augusto A. Enteiche \\ \textit{Soil-Cement: Its Use in Building}. 1964.
\end{flushright}


\begin{flushright}
\small{
\textit{
``Contemporary stabilized rammed earth (SRE) draws upon traditional rammed earth (RE) methods and materials, often incorporating reinforcing steel and rigid insulation, enhancing the structural and energy performance of the walls while satisfying building codes. SRE structures are typically engineered by licensed Structural Engineers using the Concrete Building Code or the Masonry Building Code."}} \\ --- Bly Windstorm and Arno Schmidt. \\ \textit{A Report of Contemporary Rammed-Earth Construction and Research in North America}. 2013.
\end{flushright}


Generally, the above quotes represent modern/contemporary models concerning rammed-earth's function as a building material, physically, and with regards to its respective building culture. Endemic to both rationalizations is [what is seen to be] a destructive reduction of vital qualities (or non-qualities) of rammed-earth building, e.g. ``resistant properties" and ``energy performance \cite{MOECONVERGENCE}."  McLuhan's wisdom still applies, the rammed-earth v2.0-medium conveys the message that, as Bruce Sterling noted (\textit{Shaping Things}, 2005), the model is the message. This is to say that rammed-earth \textit{in rem} does not necessarily possess the property of sustainability, instead, the veritable sustainability of rammed-earth is a phenomenon emerging from the inanimate and animal collectives involved in rammed-earth building, i.e. the model.

\subsection{How Much Does Your Building Weigh?}

 massiveness: opportunity and challenge: transportation and stabilization

 emergy and intelligent standardization
